---
title: Hold That Value
subtitle: "The Convenience of Global Variables Without All the Danger"
author: "R. Mark Sharp"
date: "5/11/2020"
output: html_document
---

```{r setup, include = FALSE, eval = TRUE}
knitr::opts_chunk$set(echo = TRUE)
if (!require('RWordPress')) {
  devtools::install_github(c("duncantl/XMLRPC", "duncantl/RWordPress"))
}
library(knitr)
library(RWordPress)
options(WordpressLogin = c(msharp2013 = "Ilikewordpress"), 
        WordpressURL = "http://rmsharp.me/xmlrpc.php")

```
```{r knit2wp, eval = FALSE, echo = FALSE}
# Optional: If you want to see all of the knit2wp arguments that are possible, 
# run this line.
knit2wp

# If need be, set your working directory to the location where you stored the Rmd file. 

# Tell knitr to create the html code and upload it to your WordPress site
setwd("~/Documents/Projects/Active_Projects/Training/2020/blogs/vignettes")
knit2wp(
  "hold_that_value.Rmd", 
  title = "Hold That Value: Using Exported Functions Like Global Variables",
  publish = FALSE)

```

## Global Variables Are Convenient

I wanted a global variable to serve as a flag that would keep track of state
as I parsed some text. 
I could not use a local variable inside the parsing function that was being
called repeatedly and for several other use specific reasons I wanted a simple
flag with a value of `TRUE` or `FALSE`. 

I copied the idea from a short article, [A Little R Counter](
https://www.r-bloggers.com/a-little-r-counter/)
on www.r-bloggers.com.


```{r rblogger-code, echo = TRUE, eval = TRUE}
CreateCounter <- function(curr.count) {
  list(
    increment = function(amount) {
      curr.count <<- curr.count + amount
    },
    value = function() {
      return(curr.count)
    }
  )
}

```

This lets you do the following:

```{r counter-working, echo = TRUE, eval = TRUE}
counter <- CreateCounter(10)
counter$increment(9)
counter$value()

```
I wanted to create something only slightly different. 
I did not need to count but I did want to retain state.

```{r make-true-false, echo = TRUE, eval = TRUE}
make_true_false_flag <- function(set_value) {
  if (is.null(set_value) | !"logical" %in% class(set_value))
    stop("Must provide seed value of TRUE or FALSE")
  
  list(
    set = function(value) {
      set_value <<- value[[1]] # protection against > 1 value
    },
    value = function() {
      return(set_value)
    }
  )
}

```
This allows the following simple uses:

```{r make-true-false-examples, echo = TRUE, eval = TRUE}
# Create function list with `value` defaulted to TRUE
tf_flag <- make_true_false_flag(TRUE)
tf_flag$value()
# reset `value` 
tf_flag$set(FALSE)
tf_flag$value()
# Create function list with `value` defaulted to FALSE
tf_flag <- make_true_false_flag(FALSE)
tf_flag$value()
# reset `value` 
tf_flag$set(TRUE)
tf_flag$value()
```

Of course, as soon as I used the function, it felt a bit of a waste to write
a function with such limited utility when only slightly more code an no
more complexity, I could hold and set any state I wanted.

The following code accomplishes that.

```{r make-any-state-holder, echo = TRUE}
make_flag <- function(set_value, flags = c(TRUE, FALSE)) {
  if (is.null(set_value) | !is.element(set_value, flags))
    stop("Must provide seed value from 'flags' vector")
  
  list(
    set = function(value) {
      set_value <<- value[[1]] # protection against > 1 value
    },
    value = function() {
      return(set_value)
    }
  )
}

```

Of course this can serve to make a `TRUE`, `FALSE` flag holder as demonstrated
below.

```{r limiting-to-true-false, echo = TRUE}
# Create function list with `value` defaulted to TRUE
tf_flag <- make_flag(TRUE, c(TRUE, FALSE))
tf_flag$value()
# reset `value` 
tf_flag$set(FALSE)
tf_flag$value()
# Create function list with `value` defaulted to FALSE
tf_flag <-make_flag(TRUE, c(TRUE, FALSE))
tf_flag$value()
# reset `value` 
tf_flag$set(TRUE)
tf_flag$value()

```

However, the slight increase in complexity allows you to store, reset, and 
retrieve from any vector you want.
```{r examples-any-flag, echo = TRUE}
#' # Create function list with `value` defaulted to TRUE
tf_flag <- make_flag("defined", c("defined", "pending", "completed"))
tf_flag$value()
# reset `value` 
tf_flag$set("pending")
tf_flag$value()
# Create function list with `value` defaulted to "S"
letter_flag <- make_flag("S", LETTERS)
letter_flag$value()
# reset `value` 
letter_flag$set("X")
letter_flag$value()
# Of course numeric or integer values are allowed
integer_holder <- make_flag(4, 1:10)
integer_holder$value()
# etc.

```


